---
title: "Assignment 2: Coding Basics"
author: "Srishti Mutha"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 
sequence1 <- seq(1,100,4) #using sequence code and assigning it the name sequence1
sequence1 


#2. 
mean(sequence1) #generating mean using code mean(sequence1)
median(sequence1) #generating median using code median(sequence1)

#3. 
mean(sequence1)>median(sequence1) #determining if mean is greater than median
```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
students <- c('Cherry','Sam','Joe','Fin') # it is a string vector
scores<-c(80,88,90,20) # it is integar vector
pass<-ifelse(scores>50,'Pass','fail')
students
scores
pass

table<-data.frame(students, scores, pass) #generating the table using cod
table
```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:Data frame can have different classes of data while a matrix can only have a single class of data.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}
scores_and_print<-function(x)
{
  ifelse(x>50, 'TRUE','FALSE')
}
scores_and_print(scores)

#Tried the 'if' and 'else' statements but this error occured - Error in if (x > 50) { : the condition has length > 1

#scores_result<-function(x)

 # if(x>50) {
   # print('TRUE')
 # } else {
  #  print('FALSE')
#  }

#scores_result(scores)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer:The if/else function didnot work as it only checks one vector at a time whereas ifesle function checks each vector.
